---
title: "Homework 6"
author: "Jørgen Højlund Wibe"
output:
  pdf_document: default
  html_document: default
subtitle: Cultural Datascience 2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading data**
```{r}
library(pacman)
pacman::p_load(gapminder, tidyverse)
df <- gapminder
```

### Task 1
**Task description**

*Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. You can use the population and GDPpercapita columns for it.*

*Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.*

**Solution**
```{r}
# Defining function
find_gdp <- function(df, year=NULL, country=NULL) {
  if (!is.numeric(year)) {
    warning(stop("Year must be of type numeric"))
  }
  if (!is.character(country)) {
    warning(stop("Country must be of type character"))
  }
  if(!is.null(year)) { # if the 'year' variable is not empty...
    df <- df[df$year %in% year, ] # add the specified year to df
  }
  
  if (!is.null(country)) { # same as above
    df <- df[df$country %in% country, ] # -||-
  }
  
  gdp <- df$pop * df$gdpPercap # calculating gdp: multiplies population with GDP per capita

  new_df <- cbind(df, gdp=gdp) # combining results into one dataframe
  return(new_df)
}

# Running function, specifying values
find_gdp(df, year = c(1967, 1977, 1987, 1997, 2007, 2017), country = "Denmark" )
```



### Task 2
**Task description**

*Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and prints out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70. (Hint: remember the grepl function, and review the Control Flow tutorial (https://swcarpentry.github.io/r-novice-gapminder/07-control-flow/index.html))*

**Solution**
```{r}
countries <- unique(df$country)

for (i in countries) {
  life_expectancy <- df %>% 
    filter(country == i) %>% 
    pull(lifeExp)
  
  if ((mean(life_expectancy) < 50) && (i %in% grep("^B", countries, value = T))) {
    print(paste0(i, " starts with 'B' and has a average life expectancy less than 50!"))
  }
  
  else if (between(mean(life_expectancy), 50, 70) && (i %in% grep("^B", countries, value = T))) {
    print(paste0(i, " starts with 'B' and has an average life expectancy between 50 and 70"))
  }
  
  else if (mean(life_expectancy) > 70 && (i %in% grep("^B", countries, value = T))) {
    print(paste0(i, " starts with 'B' and has an average life expectancy more than 70!"))
  }
}
```