---
title: "Homework 3"
author: "WIBE"
date: "8/30/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Instructions: For this assignment, you need to answer a couple questions with code and then take a screenshot of your working environment.**

### 1) Use R to figure out how many elements in the vector below are greater than 2 and then tell me what their sum (of the larger than 2 elements) is.
```{r}
rooms <- c(1, 2, 4, 5, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, 9, 3, NA)

# Counting the sum of numbers greater than two excluding the NA's
length(na.omit(rooms) > 2)

sum(na.omit(rooms) > 2)
```
Answer:

### 2) What type of data is in the 'rooms' vector?
'Class' function tells us the class of a vector:
```{r}
class(rooms)
```


### 3) Installing tidyverse, loading data using read_csv()
Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.
```{r}
library(tidyverse)
interviews <- read_csv("/Users/WIBE/Desktop/CogSci/Cultural data science/Introduction course/Day 2/data/SAFI_clean.csv")
```


### 4) Challenge: If you managed to create your own Danish king dataset, use it. If not, you the one attached to this assignment (it might need to be cleaned up a bit). Load the dataset into R as a tibble. Calculate the mean() and median() duration of rule over time and find the three monarchs ruling the longest. How many days did they rule (accounting for transition year?)
```{r}
kings_and_queens <- read_csv(as_tibble("/Users/WIBE/Desktop/CogSci/Cultural data science/Introduction course/Day 2/data/Danish Monarchs - Sheet1.csv"))
```

### Calculating mean and median duration of rule over time
```{r}
median((kings_and_queens$end_reign)-(kings_and_queens$start_reign), na.rm = T)

mean((kings_and_queens$end_reign)-(kings_and_queens$start_reign), na.rm = T)
```
Mean reign length is: **47 years**
Median reign length is: **43 years**

### Finding the three monarchs ruling the longest
```{r}
kings_and_queens %>% 
  mutate(reign_length = end_reign-start_reign) %>% 
  filter(reign_length > 42) %>% 
  select(Monarch,reign_length)
```
The three longest reigning kings were

  Monarch            reign_length
  
1 Erik 7. af Pommern           43
2 Christian 4.                 60
3 Christian 9.                 43






